---
title: "LogisticRegression-A1"
author: "Sanyamc"
date: "Monday, June 29, 2015"
output: html_document
---


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}

songs <- read.csv("E:\\github\\Courses\\Analytics-Edge\\DataSets\\songs.csv")
str(songs)



```

Here's a detailed description of the variables:
.year = the year the song was released
.songtitle = the title of the song
.artistname = the name of the artist of the song
.songID and artistID = identifying variables for the song and artist
.timesignature and timesignature_confidence = a variable estimating the time signature of the song, and the confidence in the estimate
. loudness = a continuous variable indicating the average amplitude of the audio in decibels
. tempo and tempo_confidence = a variable indicating the estimated beats per minute of the song, and the confidence in the estimate
. key and key_confidence = a variable with twelve levels indicating the estimated key of the song (C, C#, . . ., B), and the confidence in the estimate
. energy = a variable that represents the overall acoustic energy of the song, using a mix of features such as loudness
.pitch = a continuous variable that indicates the pitch of the song
. timbre_0_min, timbre_0_max, timbre_1_min, timbre_1_max, . . . , timbre_11_min, and timbre_11_max = variables that indicate the minimum/maximum values over all segments for each of the twelve values in the timbre vector (resulting in 24 continuous variables)
.Top10 = a binary variable indicating whether or not the song made it to the Top 10 of the Billboard Hot 100 Chart (1 if it was in the top 10, and 0 if it was not)



Splitting the data in training and testing set:
```{r}
songsTest = subset(songs,songs$year==2010)
songsTrain = subset(songs,songs$year!=2010)

```

As expected, our outcome variable is Top10, we are trying to predict weather a song will make it to top10 or not.


```{r}
# We care only about numerical variables:
nonvars = c("year","songtitle","artistname","songID","artistID")

songsTrain = songsTrain[,!(names(songsTrain)%in%nonvars)]
songsTest = songsTest[,!(names(songsTest)%in%nonvars)]

songsLog1 = glm(Top10~ .,data=songsTrain,family="binomial")

summary(songsLog1)

# Note that we can only do the -sign for numerical variables

songsLog2 = glm(Top10~ . -loudness,data=songsTrain,family="binomial")
summary(songsLog2)


songsLog3 = glm(Top10~ . -energy,data=songsTrain,family="binomial")
summary(songsLog3)

predictTest = predict(songsLog3,type="response",newdata = songsTest)

table(songsTest$Top10,predictTest>0.45)
accuracy = 328/373




```

### Parole Dataset

male: 1 if the parolee is male, 0 if female
.race: 1 if the parolee is white, 2 otherwise
.age: the parolee's age (in years) when he or she was released from prison
.state: a code for the parolee's state. 2 is Kentucky, 3 is Louisiana, 4 is Virginia, and 1 is any other state. The three states were selected due to having a high representation in the dataset.
.time.served: the number of months the parolee served in prison (limited by the inclusion criteria to not exceed 6 months).
.max.sentence: the maximum sentence length for all charges, in months (limited by the inclusion criteria to not exceed 18 months).
.multiple.offenses: 1 if the parolee was incarcerated for multiple offenses, 0 otherwise.
.crime: a code for the parolee's main crime leading to incarceration. 2 is larceny, 3 is drug-related crime, 4 is driving-related crime, and 1 is any other crime.
.violator: 1 if the parolee violated the parole, and 0 if the parolee completed the parole without violation.


```{r}

parole <- read.csv("E:\\github\\Courses\\Analytics-Edge\\DataSets\\parole.csv")
summary(parole)

parole$state = factor(parole$state)
parole$crime = factor(parole$crime)

set.seed(144)

library(caTools)

split = sample.split(parole$violator, SplitRatio = 0.7)

train = subset(parole, split == TRUE)

test = subset(parole, split == FALSE)

vlLog = glm(violator~ .,data=train,family="binomial")

summary(vlLog)

## To compute the odds = e^coeffiecient of the variable




```

