---
title: "AnalyticsEdge-Unit1-HW"
author: "Sanyamc"
date: "Tuesday, April 14, 2015"
output: html_document
---


    Analytics Edge
========================
This is continued report/notes from Chapter 1 assignments.

Stock Markets
----------------------
Here we are looking at monthly stocks data from 5 companies, IBM, GE, Coke, PnG,Boeing from 1970-2009

Lets take a look at summary of stock prices for these companies. 



```{r}
ibm=read.csv('AnalyticsEdge/IBMStock.csv')
ge=read.csv('AnalyticsEdge/gestock.csv')
png=read.csv('AnalyticsEdge/procterGamblestock.csv')
boeing=read.csv('AnalyticsEdge/boeingstock.csv')
coke=read.csv('AnalyticsEdge/cocacolastock.csv')

ibm$Date=as.Date(ibm$Date,'%m/%d/%y')
ge$Date=as.Date(ge$Date,'%m/%d/%y')
coke$Date=as.Date(coke$Date,'%m/%d/%y')
png$Date=as.Date(png$Date,'%m/%d/%y')
boeing$Date=as.Date(boeing$Date,'%m/%d/%y')

summary(ibm)
summary(ge)
summary(coke)
summary(boeing)
summary(png)

ibm$Date[which.max(ibm$StockPrice)]

```

Checkout the following ***line plot*** below
- Note the multiple geoms
- Note the date breaks
- Note the ***vertical line***
```{r}

ggplot(aes(x=Date,y=StockPrice),data=coke)+geom_line(data=png,aes(color='PnG'))+geom_line(data=ibm,aes(color='IBM',colour='black'))+geom_line(data=boeing,aes(color='Boeing'))+geom_line(data=ge,aes(color='GE'))+geom_line(aes(color='Coke'))+scale_x_date(breaks="2 year")+geom_vline(xintercept=as.numeric(as.Date('2000-03-01')),alpha=1/4,linetype=2)

```
Lets observe the data withing 5 year period.
```{r}

ggplot(aes(x=Date,y=StockPrice),data=coke)+geom_line(data=png,aes(color='PnG'))+geom_line(data=ibm,aes(color='IBM',colour='black'))+geom_line(data=boeing,aes(color='Boeing'))+geom_line(data=ge,aes(color='GE'))+geom_line(aes(color='Coke'))+scale_x_date(limits=c(as.Date("2004-01-01"),as.Date("2006-01-01")))+geom_vline(xintercept=as.numeric(as.Date('2000-03-01')),alpha=1/4,linetype=2)+scale_y_discrete(breaks=seq(0,200,5))

```

Lets calculate the mean data of IBM by month across all years and then identify when IBM had historically higher stock prices than other months.

```{r}
tapply(ibm$StockPrice,months(ibm$Date),mean)

tapply(ge$StockPrice,months(ge$Date),mean)

tapply(png$StockPrice,months(png$Date),mean)

```

