---
title: "Unit1-EmploymentStats"
author: "Sanyamc"
date: "Tuesday, April 14, 2015"
output: html_document
---
```{r}
cps<-read.csv('AnalyticsEdge/CPSData.csv')
mac<-read.csv('AnalyticsEdge/MetroAreaCodes.csv')
cc<-read.csv('AnalyticsEdge/CountryCodes.csv')

summary(cps)
str(cps)

```

Lets sort the regions by number of Interviews in the region and state

```{r}

sort(table(cps$Region))
sort(table(cps$State))

```

What proportions of interviewees are citizens of US

```{r}
table(cps$Citizenship)
```

Whats the distribution of Race w.r.t Hispanic

```{r}
table(cps$Race,table$Hispanic)
```

Evaluate Missing values
- Checkout missing values based on the region, sex, Age, Citizenship, or none.
```{r}
table(cps$Region,is.na(cps$Married))
table(cps$Sex,is.na(cps$Married))

table(cps$Age,is.na(cps$Married))
```

- Lets explore another variable, known as MetroAreaCode and find the interviwees living in non-metro Area
```{r}
table(cps$State,is.na(cps$MetroAreaCode))
table(cps$Region,is.na(cps$MetroAreaCode))
```

Now we have been calculating the proportions using the table command and then manually computing the values.
Instead we can use the ***mean*** of 0 and 1 and that would give use the  right value.

Follow along to understand :)
```{r}
sort(tapply(is.na(cps$MetroAreaCode),cps$State,mean))
```

Merging data from other dictionaries
---------------------------------------------------
```{r}
cps=merge(cps,mac,by.x="MetroAreaCode",by.y="Code",all.x=TRUE)
summary(cps)

```

Checkout the new MetroArea Columns above. In the function above all.x=TRUE means keeps the rows even if merge don't match. In other words, its a left outer join instead of inner join.

Lets use tapply method to calcuate the proportions and use this new column
- Identify the interviewers from each area
- Identify the proportions of MetroArea with Hispanic population
- Identify the proportions of Asian distribution 
- Identify the area which has no high school education

```{r}
table(cps$MetroArea)

sort(tapply(cps$Hispanic,cps$MetroArea,mean))

tapply(cps$Race=='Asian',cps$MetroArea,mean)

sort(tapply(cps$Education=="No high school diploma",cps$MetroArea,mean,na.rm=TRUE))

```

Now lets merge the country codes data
- What proportions of interviewees from NY area have country of birth that is not USA
- Which metro area has largest number of interviewees with a country of birth in India.
- Use tapply, when I can, we can easily sort it. Remember to do na.rm==TRUE for that.

```{r}
cps=merge(cps,cc,by.x="CountryOfBirthCode",by.y="Code",all.x=TRUE)
summary(cps)

sort(tapply(cps$Country!="United States",cps$MetroArea,mean,na.rm=TRUE))

sort(tapply(cps$Country=="India",cps$MetroArea,sum,na.rm=TRUE))

```




