---
title: "Assignment"
author: "Sanyamc"
date: "Wednesday, March 30, 2016"
output: html_document
---

# Clear Workspace
rm(list=ls())
# Clear Console:
cat("\014")


Review the patterns described in DataScience01a.R and use R to get the Indian Liver Patient Dataset
from the UCI machine learning repository


```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/00225/Indian%20Liver%20Patient%20Dataset%20(ILPD).csv"

ILPD <-read.csv(url,header=FALSE,stringsAsFactors=FALSE)
```

- Gather 11 column names from http://archive.ics.uci.edu/ml/datasets/ILPD+(Indian+Liver+Patient+Dataset) 
and add headers to the data frame.
- Observe first 6 rows of data frame.

```{r}
a.headers = c("Age","Gender","TB","DB","Alkphos","Sgpt","Sgot","TP","ALB","A/G","Selector")
names(ILPD) = a.headers
head(ILPD)

```

Review the patterns described in DataScience01a.R. Write code to determine the mean, median,
and standard deviation (sd) of each column and present their values in the console. Some
calculations may fail. Where applicable, fix the failures by using na.rm = TRUE. Type ?median to see
how.


```{r}
summary(ILPD)
sapply(ILPD,mean,na.rm=TRUE)
sapply(ILPD,median,na.rm=TRUE)
sapply(ILPD,sd,na.rm=TRUE)

```


