---
title: "LinearRegression-Basketball"
author: "Sanyamc"
date: "Saturday, April 18, 2015"
output: html_document
---

Moneyball --> Basketball
============================

This is part of recitation for Linear regression where we are going to explore Basketball team's data sets and apply moneyball techniques there.

Lets load both the available datasets basketball train and basketball test in R 

```{r}
nba_train<-read.csv('E:\\github\\Courses\\Analytics-Edge\\DataSets\\nba_train.csv')
nba_test<-read.csv("E:\\github\\Courses\\Analytics-Edge\\DataSets\\nba_test.csv")

str(nba_test)
```
As we can see there are tons of variables here.

Our goal is to find out how many wins a team needs to win in order to make the playoffs.

```{r}
table(nba_train$W,nba_train$Playoff)

```
As we can see from the table command, a team that wins about 42 games is gets a playoffs spot. i.e. in the past team that wins 42 games have been to playoffs 29 times and not in playoffs around 8 times.

Just like the runs difference, we are going to use the points difference and see the linear relation between the two.

```{r}
nba_train$pd<-nba_train$PTS-nba_train$oppPTS
library('ggplot2')
ggplot(aes(x=W,y=pd),data=nba_train)+geom_point()
```

Obviously there is a **strong** linear relation between points diff and wins. Lets create a linear model
```{r}
model1<-lm(W~pd,nba_train)
summary(model1)
```

As we can see that this a good model as R2 is close to 1.
Our linear regression eq is W = 41+.03259(pd)

In terms of points difference for a team making playoffs(winning 42 games) team needs to have 1/.0326 pointsdiff which comes out to be around 31.

Now the PD largly depends on PS and PC
```{r}
model1<-lm(PTS~X2PA+X3PA+FTA+ORB+DRB+AST+STL+BLK+TOV,data=nba_train)
summary(model1)
```

Now lets remove the variables which are less sig. From the above table that would be TOV (p very high)

```{r}
model2<-lm(PTS~X2PA+X3PA+FTA+ORB+DRB+AST+STL+BLK,data=nba_train)
summary(model2)
model3<-lm(PTS~X2PA+X3PA+FTA+ORB+DRB+AST+STL,data=nba_train)
summary(model3)
```

Although R^2 remained the same but other measures like Root mean squared error could have increased so check them out before sticking with the new model.

Now using this model3 lets make prediction for 2012-2013 season in the test data set.

```{r}
outcome<-predict(model3,nba_test)
summary(outcome)
SSE<-sum((nba_test$PTS-outcome)^2)
SST<-sum((nba_test$PTS-mean(nba_train$PTS))^2)
R2<-1-(SSE/SST)
RMSE<-sqrt(SSE/length(nba_test$PTS))
```
As we can see that our R2 is pretty high and our RMSE but not too bad making an avg. error of about 196 points 

